---
editor_options:
  markdown:
    wrap: 72
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

## Work breakdown

-   *Vitalii Sharaiev*: conditional probability function, predict,
    score.

-   *Olga Havryliuk*: all additional function, predict.

-   *Mariia Ivanchenko*: prepare data, vizualization, predict.

## Introduction

**Bayes formula:**
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

## Data description

This data set contains data of American news: a headline and an abstract
of the article. Each piece of news is classified as fake or credible. We
will classify the news from test.csv as credible or fake.

```{r}
# list of libraries that we used
library(dplyr)
library(tidyr)
library(tokenizers)
library(tidytext) 
library(readr)
library(ggplot2)
library(wordcloud2)
library(caret) 
```

## Outline of the work

1.  **Data pre-processing** (includes removing punctuation marks and
    stop words, representing each message as a bag-of-words)
2.  **Data visualization** (it's time to plot your data!)
3.  **Classifier implementation** (using the training set, calculate all
    the conditional probabilities in formula (1) and then use those to
    predict classes for messages in the testing set)
4.  **Measurements of effectiveness of your classifier** (accuracy,
    precision and recall curves, F1 score metric etc)
5.  **Conclusions**

## Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

### Reading csv files

```{r}
test_path <- "test.csv"
train_path <- "train.csv"
stop_words <- "srop_words.txt"
cleanTrainPath <- "cleaned.csv"
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)
test <-  read.csv(file = test_path, stringsAsFactors = FALSE)
clean <- read.csv(file = cleanTrainPath, stringsAsFactors = FALSE)
```

### Preparing our training file in order to use it during training our model,

Before feeding the data into the classifier, the following preprocessing
steps were taken:

-   Tokenizing the text data into individual words

-   Removing punctuation, numbers, and other non-word characters

-   Converting text to lowercase for uniformity.

```{r}
clean_text <- function(stop_words_file, text_file_path){
  text <- read.csv(file = train_path, stringsAsFactors = FALSE)
  stop_words <- unlist(strsplit(read_file("stop_words.txt"), "\n"))
  pattern <- paste(stop_words, collapse = "|")
  for (i in 1:nrow(text)) {
    text$Body[i] <- gsub(paste0("\\b(", pattern, ")\\b"), "", text$Body[i], ignore.case = TRUE) 
    text$Body[i] <- gsub(pattern = "([[:punct:]])", "", text$Body[i], ignore.case = TRUE)
    text$Body[i] <- gsub(pattern = "[0-9]", "", text$Body[i], ignore.case = TRUE)
  }
  write.csv(text, "cleaned.csv")
}
cleanTrain <- clean_text(stop_words, train_path)
```

## Classifier implementation

**1. Class Fields:**\
- **data:** This field holds the entirety of our dataset\
- **labels:** An array comprising two labels: 'fake' and 'credible'\
- **fake_article_probability:** Probability that a news article is fake\
- **credible_article_probability:** Probability that a news article is
credible\
- **all_vocabulary:** A data frame that encapsulates all word
occurrences across articles\
- **fake_words_dictionary:** A data frame detailing occurrences of words
in fake news\
- **credible_words_dictionary:** A data frame detailing occurrences of
words in credible news

**2. Class Functions:**\
- **fit(X, Y):** Processes data and labels to set up the model\
- **probability_of_one_word(word, label):** Calculates how likely a word
fits the 'credible' or 'fake' category. - **predict(message):** Uses the
Bayes formula to determine the credibility of a message\
- **score(X_test):** Produces predictions to help assess the model's
performance.

```{r}

naiveBayes <- setRefClass("naiveBayes",
                          
                          # Class Fields
                          fields = list(
                            data = "data.frame",
                            dataForTest = 'data.frame',
                            dataSetWithPrediction = "data.frame",
                            lables = "vector",
                            cleaned_data = "data.frame",
                            credible_words = "list",
                            fake_words = "list",
                            laplaceConst = "numeric",
                            total_credible_words = "numeric",
                            total_fake_words = "numeric",
                            total_fake_class = "numeric",
                            total_credible_class = "numeric",
                            
                            cred_cond_prob = "list",
                            fake_cond_prob = "list",
                            words = "numeric",
                            
                            accuracy = "numeric",
                            recall = "numeric",
                            f1_score = "numeric",
                            precision = "numeric"
                          ),
                          
                          methods = list(
                            
                            # Transforms the main data to get 'Label' and 'splitted' columns
                            text_dataframe = function(main) {
                              tidy_text <- main %>%
                                unnest_tokens(splitted, Body, token="words")
                              
                              return(tidy_text[, c("Label", "splitted")])
                            },
                            
                            # Counts unique word occurrences in the dataset
                            words_occurrences = function(main) {
                              tidy_text <- main %>%
                                unnest_tokens(splitted, Body, token="words")
                              distinct_words <- unique(tidy_text$splitted)
                              return(length(distinct_words))
                            },
                            
                            # Counts occurrences of words labeled as 'credible'
                            credible_count = function(df) {  
                              credible_word_counts_df <- df %>%
                                dplyr::filter(Label == lables[1]) %>%
                                count(splitted, sort = TRUE)
                              result_dict <- as.list(setNames(credible_word_counts_df$n, credible_word_counts_df$splitted))
                              return(result_dict)
                            },
                            
                            # Counts occurrences of words labeled as 'fake'
                            fake_count = function(df) {   
                              fake_word_counts_df <- df %>%     
                                dplyr::filter(Label == lables[2]) %>%     
                                count(splitted, sort = TRUE)      
                              result_dict <- as.list(setNames(fake_word_counts_df$n, fake_word_counts_df$splitted))      
                              return(result_dict) 
                            }, 
                            
                            # Returns a list of word counts from a given message
                            words_summary = function(text) {   
                              tidy_text <- unnest_tokens(data.frame(text=text), 'splitted', 'text', token="words")   
                              word_counts <- tidy_text %>% count(splitted, sort=TRUE)   
                              result_dict <- as.list(setNames(word_counts$n, word_counts$splitted))      
                              return(result_dict)
                            },
                            
                            # Processes data to initialize the model
                            fit = function() {
                              
                              cleaned_data <<- text_dataframe(data)
                              words <<- words_occurrences(data)
                              
                              credible_words <<- credible_count(cleaned_data)
                              fake_words <<- fake_count(cleaned_data)
                              
                              total_credible_words <<- sum(unlist(credible_words))
                              total_fake_words <<- sum(unlist(fake_words))
                              
                              total_fake_class <<- sum(data$Label == lables[2])
                              total_credible_class <<- sum(data$Label == lables[1])
                            },
                            
                            # Computes the conditional probability of a word given a class (credible or fake)
                            conditionalProbability = function(word, specificClass) {
                              if (specificClass == 1) {
                                totalNumberOfOccurenceInClassOfWord <- ifelse(is.null(credible_words[[word]]), 0, credible_words[[word]])
                                totalClassWords <- total_credible_words
                              } else {
                                totalNumberOfOccurenceInClassOfWord <- ifelse(is.null(fake_words[[word]]), 0, fake_words[[word]])
                                totalClassWords <- total_fake_words
                              } 
                              numerator <- totalNumberOfOccurenceInClassOfWord + 1
                              denominator <- totalClassWords + words
                              
                              probability <- numerator / denominator
                              
                              return(probability)
                            },
                            
                            
                            # Predicts the credibility of a message
                            predict = function(message) {
                              message <- gsub('[[:punct:]0-9 ]+', ' ', message)
                              message <- tolower(message)
                              
                              fakeArticle <- (total_fake_class) / (total_fake_class + total_credible_class)
                              credibleArticle <- 1 - fakeArticle
                              
                              message_summary <- words_summary(message)
                              
                              for (key in names(message_summary)) {
                                fakeArticle <-fakeArticle * conditionalProbability(key, 0)  * 10000 
                                credibleArticle <- credibleArticle * conditionalProbability(key, 1)  * 10000
                              }
                              if (fakeArticle > credibleArticle) {
                                return(lables[2])
                              } else {
                                return(lables[1])
                              }
                              
                            },
                            
                            # Adds a prediction column to the dataset
                            createDataWithPredictions = function() {
                              dataForTest["Prediction"] <<- apply(dataForTest['Body'],1, FUN = .self$predict)
                              dataSetWithPrediction <<- dataForTest
                            },
                            
                            
                            # Scores the model
                            score = function(){
                              dataSetWithPrediction <<- createDataWithPredictions()
                              actual <- dataSetWithPrediction$Label
                              predicted <- dataSetWithPrediction$Prediction
                              
                              TP <- sum(actual == lables[1] & predicted == lables[1])
                              FP <- sum(actual == lables[2] & predicted == lables[1])
                              TN <- sum(actual == lables[2] & predicted == lables[2])
                              FN <- sum(actual == lables[1] & predicted == lables[2])
                              
                              accuracy <<- (TP + TN) / (TP + FP + TN + FN)
                              recall <<- TP / (TP + FN)
                              precision <<- TP / (TP + FP)
                              f1_score <<- 2 * (precision * recall) / (precision + recall)
                              cat("Accuracy = ", accuracy, "\n")
                              cat("Recall = ", recall, "\n")
                              cat("Precision = ", precision, "\n")
                              cat("F1 Score = ", f1_score)
                            }
                          )
)

lables <- unique(train$Label)
model <- naiveBayes$new(data=clean, dataForTest = test, lables=lables)
model$fit()
model$score()

```

# Visualization

## First of all we decided to visualise the distribution of news labels in the dataset

The chart provides a visual representation of the proportion of "Fake"
news relative to "Credible" news. This visualization helps in
understanding the balance (or imbalance) of the dataset in terms of news
credibility. An imbalanced dataset could influence the training and
performance of the machine learning model.

```{r}
viz_lables_before = function() {
  slices <- c(model$total_fake_class, model$total_credible_class) # total_fake_class, total_credible_class
  lbls <- paste(c("Fake: ", "Credible: "), slices)
  pie(slices, labels = lbls, main = "Chart of Labels")
}

viz_lables_before()
```

## Visualization of top words in fake news

This word cloud visualizes the words most frequently associated with
"Fake" news.

```{r}

viz_fake_words = function() {
  wordcloud2(data = data.frame(word = names(model$fake_words), freq = unlist(model$fake_words)), color = "red")
}
viz_fake_words()

```

## Visualization of top words in credible news

This word cloud visualizes the words most frequently associated with
"Credible" news.

```{r}
viz_credible_words = function() {
  wordcloud2(data = data.frame(word = names(model$credible_words), freq = unlist(model$credible_words)), color = "green")
}

viz_credible_words()
```

### This bar plot shows the number of credible messages predicted as credible in green and the number of credible messages predicted as fake in red.

```{r}
actual <- model$dataSetWithPrediction$Label
predicted <- model$dataSetWithPrediction$Prediction

cm <- confusionMatrix(as.factor(predicted), as.factor(actual), positive = lables[1])
print(cm)
actual_credible <- sum(actual == lables[1])
predicted_credible <- sum(predicted == lables[1] & actual == lables[1])
predicted_as_fake <- actual_credible - predicted_credible

values <- c(predicted_credible, predicted_as_fake)
labels <- c("Predicted Credible", "Predicted as Fake")

barplot(values, names.arg = labels, main = "Credible Messages Predictions", col = c("green", "red"), ylim = c(0, max(values) + 10))
legend("topright", c("Predicted Credible", "Predicted as Fake"), fill = c("green", "red"))

```

### This bar plot shows the number of fake messages predicted as fake in red and the number of fake messages predicted as credible in green

```{r}
actual <- model$dataSetWithPrediction$Label
predicted <- model$dataSetWithPrediction$Prediction

cm <- confusionMatrix(as.factor(predicted), as.factor(actual), positive = lables[2])
print(cm)
actual_fake <- sum(actual == lables[2])
predicted_fake <- sum(predicted == lables[2] & actual == lables[2])
predicted_as_credible <- actual_fake - predicted_fake

values <- c(predicted_fake, predicted_as_credible)
labels <- c("Predicted Fake", "Predicted as Credible")

barplot(values, names.arg = labels, main = "Fake Messages Predictions", col = c("red", "green"), ylim = c(0, max(values) + 10))
legend("topright", c("Predicted Fake", "Predicted as Credible"), fill = c("red", "green"))


```

### 'Chart of Prediction' visualizes the distribution of news articles as predicted by the Naive Bayes classifier and 'Chart of Labels' visualizes the actual distribution

```{r}
viz_lables_after = function() {
  x <- 0
  y <- 0
  for(i in model$dataSetWithPrediction$Prediction){
    if (i == "credible"){
      x <- x +1
    } else{
      y <- y+1
    }
  }
  slices <- c(y, x) # total_fake_class, total_credible_class
  lbls <- paste(c("Fake: ", "Credible: "), slices)
  pie(slices, labels = lbls, main = "Chart of Prediction")
}

viz_lables_aftL = function() {
  x <- 0
  y <- 0
  for(i in model$dataSetWithPrediction$Label){
    if (i == "credible"){
      x <- x +1
    } else{
      y <- y+1
    }
  }
  slices <- c(y, x) # total_fake_class, total_credible_class
  lbls <- paste(c("Fake: ", "Credible: "), slices)
  pie(slices, labels = lbls, main = "Chart of Labels(true)")
}

viz_lables_after()

viz_lables_aftL()
```

# Accuracy **estimation**

Estimates accuracy of the model based on the given DF. First of all, we
calculate True Positive (tp), False Negative (fn) and False Positive
(fp) values in score method of naiveBayes. To calculate accuracy we use
formula (tp / (tp + fp)) (\~95,01%) Sensitivity is calculated by the
following formula (tp / (tp + fn)) (\~98,64) F1 measure is calculated
given ( \* (accurency \* sensitivity) / (accurency + sensitivity))
(\~96,79%)

```{r}
score = function(){
    dataSetWithPrediction <<- createDataWithPredictions()
    actual <- dataSetWithPrediction$Label
    predicted <- dataSetWithPrediction$Prediction
    
    TP <- sum(actual == lables[1] & predicted == lables[1])
    FP <- sum(actual == lables[2] & predicted == lables[1])
    TN <- sum(actual == lables[2] & predicted == lables[2])
    FN <- sum(actual == lables[1] & predicted == lables[2])
    
    accuracy <<- (TP + TN) / (TP + FP + TN + FN)
    recall <<- TP / (TP + FN)
    precision <<- TP / (TP + FP)
    f1_score <<- 2 * (precision * recall) / (precision + recall)
  }
```

## Conclusions

#### Implementation method

At first we had to prepare our data. The data is cleaned by removing
punctuation, numbers, and stop words. The result is saved in
*cleaned.csv* and all further training we perform using this cleaned
file.

Then we had to implement The Naive Bayes Classifier. At first we have to
create bag-of-words in our *fit* function, we set all necessary
attributes there. Then we have *predict* function, which calculates
probability that word belongs to "fake" or "credible" class, calculate
words frequencies in message and estimate it is "fake" or "credible"
according to conditional probability. Last step is to score our results.

Also, visualization. They are provided to understand the data
distribution, word frequencies, and predictions.

The model's accuracy, sensitivity, and F1 score are calculated to assess
its performance.

#### Pros&Cons

*Pros*

-   The Naive Bayes Classifier is straightforward and easy to
    understand, making it suitable for text classification tasks.

-   It can handle large datasets efficiently due to its simplicity.

*Cons*

-   The method assumes that words are conditionally independent, which
    may not hold true in some real-world scenarios.

-   Naive Bayes may not capture complex relationships between words and
    may not perform as well as more sophisticated models like deep
    learning approaches.
